---
title: "Text Analysis and Spatial Data for Economists"
subtitle: "Assignment 1"
author: zitian wang
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
  toc: yes
toc_float: true
toc_collapsed: true
--- 

```{r setup, include=T , message = F}
# Load first the libraries you will need
library(sf)
library(readr)
library(ggplot2)
library(dplyr)
library(leaflet)
library(RColorBrewer)
```


# Spatial operations

For this exercise, we will need first to import the file `swissBOUNDARIES3D_1_5_TLM_HOHEITSGEBIET.shp` containing the municipality boundary (do not use the file `swissBOUNDARIES3D_1_5_TLM_HOHEITSGRENZE.shp`) and the `.csv` file on housing price we used in the last exercise session.

To avoid issues with this exercise, you should remove the `Z` geometry (altitude of the point) of your municipality boundary object using the function `st_zm(your municipality object, drop = TRUE)`. Do not forget to save this as a new object.
```{r}
municipality <- st_read("/Users/zitian/Downloads/swissboundaries3d_2025-01_2056_5728.shp/swissBOUNDARIES3D_1_5_TLM_HOHEITSGEBIET.shp")
housing_data <- read_csv("/Users/zitian/Downloads/hp_small.csv")
municipality <- st_zm(municipality, drop = TRUE)
```

## Simplification

The municipality boundary contains a detailed (hence precise) geometry but uses a lot of memory such that you might want to simplify it first. 

In also contains several variables that might not be directly of use.

* Look at the object size using `object.size()`
* Look at the structure of the object. Could you guess which attributes might not be useful? 
  + select only the attributes related to the name of the municipality ('NAME'), its population ('EINWOHNERZ'), area ('GEM_FLAECH'), identification number ('BFS_NUMMER') and the canton ('KANTONSNUM').


```{r}
print(object.size(municipality), units = "Mb")
str(municipality)
municipality_select <- municipality[, c("NAME", "EINWOHNERZ", "GEM_FLAECH", "BFS_NUMMER", "KANTONSNUM")]
```


* Simplify the geometry using your preferred simplification method.
  + Check the CRS of the data first and the units of measure.
  + Save your preferred result in a new object (e.g. 'mun_simple')


```{r}
print(st_crs(municipality_select))

mun_simple <- st_simplify(municipality_select, dTolerance = 50)

# Visualize overlay of simplified (green) and original (red dashed) geometries
ggplot() +
  geom_sf(data = mun_simple, fill = NA, color = "green", size = 0.5) +
  geom_sf(data = municipality_select, fill = NA, color = "red", size = 1.5, linetype = "dashed", alpha = 0.8) +
  labs(title = "Overlay of Simplified (Green) and Original (Red) Geometries") +
  theme_minimal()

```

## Spatial join

We would like now to compute the average housing price for all Swiss municipalities. To avoid outliers, we prefer to select only municipalities with at least 5 postings. 

* Create a logical vector that identifies all municipalities with at least 5 postings.
  + What should you do first?
  + Which municipality object should you use, the original or the simplified one? Why?

```{r}
# Use CSV field 'bfscode1' to represent municipality identifier
posting_counts <- table(housing_data$bfscode1)
municipalities_with5 <- names(posting_counts)[posting_counts >= 5]
print(municipalities_with5)
```

## Spatial subsetting

* Create an object containing only municipalities with at least 5 postings.
```{r}
# Municipality data uses BFS_NUMMER while CSV uses bfscode1.
mun5 <- mun_simple[mun_simple$BFS_NUMMER %in% municipalities_with5, ]
```
## Spatial aggregation

* Compute the annual rental price per square meter (a_brutm_m2) per municipality (use the variable BFS_NUMMER to identify them) with at least 5 observations. First convert the variable a_brutm_m2 to numeric with, e.g., the function `type.convert(x, dec=".")`.

```{r}
# Convert rental price per square meter to numeric
housing_data$a_brutm_m2 <- type.convert(housing_data$a_brutm_m2, dec = ".")
# Aggregate average rental price by municipality identifier (using bfscode1 from CSV)
avg_price <- aggregate(a_brutm_m2 ~ bfscode1, 
                       data = housing_data[housing_data$bfscode1 %in% municipalities_with5, ], 
                       FUN = mean)
head(avg_price)
```
* Plot these municipalities

```{r}
# Municipality data uses BFS_NUMMER while CSV uses bfscode1.
mun5 <- mun_simple[mun_simple$BFS_NUMMER %in% municipalities_with5, ]

# Merge the aggregated average price (avg_price) with the spatial object.
mun5_prices_agg <- merge(mun5, avg_price, by.x = "BFS_NUMMER", by.y = "bfscode1", all.x = TRUE)

# Transform to WGS84 for leaflet (EPSG:4326)
mun5_prices_agg_wgs <- st_transform(mun5_prices_agg, crs = 4326)

# Define a color palette based on average rental price.
pal <- colorQuantile("YlOrRd", domain = mun5_prices_agg_wgs$a_brutm_m2, n = 5)
```

```{r}
# Create an interactive leaflet map.
leaflet(data = mun5_prices_agg_wgs) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%  
  addPolygons(
    fillColor = ~pal(a_brutm_m2),  
    fillOpacity = 0.7,
    weight = 1,
    color = "darkblue",
    popup = ~paste0("<strong>Municipality: </strong>", NAME,
                    "<br><strong>BFS: </strong>", BFS_NUMMER,
                    "<br><strong>Price (per m²): </strong>", round(a_brutm_m2, 2))
  ) %>%
  setView(lng = 8.2275, lat = 46.8182, zoom = 8)
```

* Add to this object the corresponding housing prices 

```{r}
# Merge detailed housing data with spatial data (if needed for further analysis)
mun5_prices <- merge(mun5, housing_data, by.x = "BFS_NUMMER", by.y = "bfscode1", all.x = TRUE)
str(mun5_prices)
```


* Plot the average housing price per square meter 

```{r}
ggplot(mun5_prices_agg) +
  geom_sf(aes(fill = a_brutm_m2)) +
  labs(fill = "Average rental price\n(per square metre)") +
  ggtitle("Average annual rental price per square metre in urban areas") +
  theme_minimal()
```


## Distances

Select all postings in the canton of Ticino that are located within a 1km distance of a municipality border
  
* How would you approach this problem? 
* What information do you need?
* **Think carefully at how distance are computed for different geometries**
* Before doing any analysis, you should remove the `Z` geometry (altitude of the point) using the function `st_zm(your municipality object, drop = TRUE)` 

```{r}
housing_sf <- st_as_sf(housing_data, coords = c("longitude2", "latitude2"), crs = 4326)

housing_sf <- st_transform(housing_sf, st_crs(mun_simple))

mun_ti <- mun_simple[mun_simple$KANTONSNUM == 21, ]
mun_ti_clean <- st_cast(mun_ti, "MULTIPOLYGON", warn = FALSE)

ti_boundaries <- st_boundary(mun_ti_clean)
ti_boundaries_union <- st_union(ti_boundaries)

distances <- st_distance(housing_sf, ti_boundaries_union)
min_distances <- as.numeric(distances)  

print(summary(min_distances))

housing_near_border <- housing_sf[min_distances <= 1000, ]

plot(st_geometry(mun_ti_clean), main = "Properties in Ticino city centre and within 1km of the city boundary", col = "lightgreen")
plot(st_geometry(housing_near_border), add = TRUE, col = "red", pch = 20, cex=0.8)

```



